---
title: "titanic"
author: "Kingsley Ofoegbu"
date: "1/28/2020"
output: pdf_document
---

```{r}
suppressMessages(library(tidyverse))
```
## Load Data
```{r}
titanic <- read_csv("titanic.csv")
```
## Clean the data
```{r}
titanic$Survived <- as.factor(titanic$Survived)
titanic$Sex <- as.factor(titanic$Sex)
titanic$Pclass <- as.factor(titanic$Pclass)
titanic$Embarked <- as.factor(titanic$Embarked)
titanic %>%
  na.omit() -> titanic2
```

## Exploratory analysis
```{r}
ggplot(titanic2, aes(Survived, Age)) +
  geom_boxplot(stat = "boxplot")
```
```{r}
ggplot(titanic2, aes(x = Sex, fill = Survived)) + 
  theme_bw() +
  geom_bar() +
  labs(y = "Passenger Count",
       title = "Titanic Survival Rates by Sex")

```
```{r}
ggplot(titanic2, aes(x = Age, color = Survived)) +
  theme_bw() +
  geom_histogram(binwidth = 5) +
  labs(y = "Passenger Count",
       x = "Age (binwidth = 5)",
       title = "Titanic Survival Rates by Age")
```

```{r}
ggplot(titanic2, aes(x = Age, fill = Survived)) +
  theme_bw() +
  facet_wrap(Sex ~ Pclass) +
  geom_histogram(binwidth = 5) +
  labs(y = "Age",
       x = "Survived",
       title = "Titanic Survival Rates by Age, Pclass and Sex")
```

## Trying to see what factors made people survive or not
```{r}
titanic.MLR <-
  glm(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked,
      family = binomial(link = logit),
      titanic2)

summary(titanic.MLR)
```
```{r}
step(titanic.MLR, direction = "both")
```
## New model after fitting with step-wise
```{r}
titanic.MLR2 <-  glm(formula = Survived ~ Pclass + Sex + Age, family = binomial(link = logit), 
    data = titanic2)
```

## Compare models 
```{r}
selcrit <-function(lmout) {
n <- length(lmout$fit)
rsq <- summary(lmout)$r.sq
adj.rsq <- summary(lmout)$adj.r.sq 
aic <- extractAIC(lmout)[2]
bic <- extractAIC(lmout, k = log(n))[2]
press <- sum((lmout$residuals/(1 - hatvalues(lmout)))^2) 
cbind(rsq, adj.rsq, aic, bic, press)
}
```
## Model 1
```{r}
selcrit(titanic.MLR)
```
## Model 2
```{r}
selcrit(titanic.MLR2)
```
## Conclusions
- Both AIC and SBC (SBC is also known as BIC) for the 3-predictor (4-parameter) model are smaller.
- PRESS for the 7-predictor (8-parameter) model is smaller.
- R2 show both models explain a large proportion of the variation in risk of infection (Y). However, we should not use R2 for model selection since these two models have different number of parameters.
-The below measures all indicate the 4-predictor, 5-parameter model is preferred over the model using all 3 predictors 

## run an anova comparing both models
```{r}
anova(titanic.MLR, titanic.MLR2)
```

